---
title: "PROG8430-Demo-Summarize"
author: "Peiyuan"
date: "05/26/2023"
output: pdf_document
---
NOTE: Comments are included for you to help you understand the code.                      
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This section is for the basic set up.
```{r echo=FALSE}
if(!is.null(dev.list())) dev.off() #checks if any graphical devices are open, if yes, close them
cat("\014") #clearing the console and starting on a new page
rm(list=ls()) #This line removes all objects from the workspace
options(scipen=9)
```
#Loading required R packages
tidyverse for easy data manipulation and visualization
caret for easy machine learning workflow

```{r}
if(!require(MASS)){install.packages("MASS")}
library(MASS)
if(!require(ggplot2)){install.packages("ggplot2")}
library(ggplot2)
if(!require(devtools)){install.packages("devtools")}
library(devtools)
if(!require(psych)){install.packages("psych")}
library(psych)
```

#Preparing the data
We’ll use the iris dataset for predicting iris species based on the predictor variables Sepal.Length, Sepal.Width, Petal.Length, Petal.Width. LDA can be affected by the scale/unit in which predictor variables are measured. It’s generally recommended to standardize/normalize continuous predictor before the analysis.

```{r}
# Load the data
attach(iris)
str(iris)
ggplot(iris, aes(Sepal.Length, Sepal.Width)) +  geom_point(aes(color = Species))

#We will create a scatterplot for the first four numerical variables (features). 
pairs.panels(iris[1:4],
             gap = 0,
             bg = c("red", "green", "blue")[iris$Species],
             pch = 21)
```


##1. Split the data into training and test set:
```{r}
# Split the data into training (60%) and test set (40%)
set.seed(123) #set the seed for random number generation in R
ind <- sample(2, nrow(iris),
              replace = TRUE,
              prob = c(0.6, 0.4)) #randomly sampling from original data - 2 groups - training/testing
training <- iris[ind==1,]
testing <- iris[ind==2,]
```

##2. Compute LDA
```{r}
linear <- lda(Species ~., training)
linear
```
##3. Make Predictions
```{r}
predictions <- predict(linear, testing)#apply model in the testing data.
# Model accuracy
mean(predictions$class==testing$Species)
```

##4. Visulize linear discrimants
```{r}
lda_plot <- cbind(training, predict(linear)$x)
ggplot(lda_plot, aes(LD1, LD2)) + geom_point(aes(color = Species))

```


